---
title: Secondary Malignancies
author: "[Benjamin Chan](http://careers.stackoverflow.com/benjaminchan)"
date: "`r Sys.time()`"
output: 
  html_document: 
    keep_md: yes
    number_sections: yes
    toc: yes
---

`source` the R script, `make.R` to generate the project document.

```
> source("make.R")
```


# Load data

Load the data from the Box folder.

```{r}
setwd("~/GitHub repositories/AEAfterBreastCaACT/SecondaryMalignancies")
list.files()
library(openxlsx)
library(data.table)
pathBox <- "C:/Users/chanb/Box Sync/NCI Systematic Review - Secondary Malignancies"
f <- file.path(pathBox, "SecondaryMalignancies_Data_BG_3_4.xlsx")
sheets <- getSheetNames(f)
sheets
D <- read.xlsx(f, sheet = sheets[1], startRow = 3, colNames = FALSE)
D <- data.table(D)
oldnames <- names(D)
newnames <- c("id",
              "authorYear",
              "trial",
              "quality",
              "arm",
              "nRandomized",
              "nITT",
              # "pctCompletingTx",
              "medianFU",
              "regimen",
              "anthracyclineType",
              "anthracyclineTotalDose",
              "anthracyclineDuration",
              "anthracyclineCourses",
              "cyclophosphamideDose",
              "cyclophosphamideDuration",
              "cyclophosphamideCourses",
              "taxaneType",
              "taxaneTotalDose",
              "taxaneDuration",
              "taxaneCourses",
              "fluoroucilTotalDose",
              "fluoroucilDuration",
              "fluoroucilCourses",
              "otherTxDetails",
              "malAML",
              "malMDS",
              "malAMLOrMDS",
              "malNonBreastSolid",
              "malNonBreastSolidType",
              "malOtherBloodCancers",
              "malSMRelatedDeaths",
              "malSecondPrimary",
              "NOTES")
setnames(D, oldnames, newnames[1:33])
```

# Tidy data

Create a `rowid` variable.
This will be handy later.

```{r}
D[, rowid := .I]
```

Input missing data for Bergh (2000); row was split.

```{r}
col <- grep("^cyclophosphamide", names(D), invert = TRUE)
D[11, col] <- D[10, col, with = FALSE]
# D[11, ]
# D[10, ]
```

Fix trial column for Romond (2005).

```{r}
D[grep("Romond", authorYear), trial := gsub("&N", "& N", trial)]
```

Make `medianFU` numeric.

```{r}
D <- D[grep("-", medianFU, invert = TRUE), medianFUNum := round(as.numeric(medianFU), digits = 2)]
x <- D[grep("-", medianFU), medianFU]
x <- as.numeric(unlist(strsplit(x, "-")))
x1 <- x[seq(1, length(x), 2)]
x2 <- x[seq(2, length(x), 2)]
xbar <- rowMeans(cbind(x1, x2))
D <- D[grep("-", medianFU), medianFUNum := xbar]
D <- D[,
       `:=` (medianFUChar = medianFU,
             medianFU = medianFUNum,
             medianFUNum = NULL)]
unique(D[, .(medianFU, medianFUChar)])
D <- D[, medianFUChar := NULL]
```

Recode mal* values of `NR` and `-` to `NA`.

```{r}
recode <- function(x) {
    missval <- c("-", "NR", " ")
    vec <- D[, get(x)]
    vec[vec %in% missval] <- NA
    vec
}
col <- grep("^mal", names(D), value = TRUE)
for (i in 1:length(col)) {
    D[, col[i]] <- recode(col[i])
}
```

Substitute non-numeric text in mal* values.

```{r}
substitute <- function(x) {
    vec <- D[, get(x)]
    vec <- gsub("[[:alpha:]]", "", vec)
    vec
}
col <- grep("^mal", names(D), value = TRUE)
for (i in 1:4) {
    # Only substitute a subset of mal* columns
    D[, col[i]] <- substitute(col[i])
}
```

Convert mal* values to numeric.

```{r}
for (i in 1:4) {
    # Only convert a subset of mal* columns
    D <- D[, col[i] := as.numeric(get(col[i]))]
}
```

The `malAMLOrMDS` column, as entered, captures number of AML or MDS malignancies if the study reported them grouped as opposed to separately.
If the study reported them as separately, then these counts are captures in the `malAML` and `malMDS` columns.
I.e., the `malAML`, `malMDS`, and `malAMLOrMDS` columns are mutually exclusive counts.
We want to create another column, `malAMLOrMDSTotal`, to be non-mutually exclusive from `malAML`, `malMDS`, and `malAMLOrMDS`.
If none of the `malAML`, `malMDS`, and `malAMLOrMDS` values are populated, then set `malAMLOrMDSTotal` to `NA`, also.
NOTE: Couldn't get this to work: [http://stackoverflow.com/a/16513949/1427069](http://stackoverflow.com/a/16513949/1427069).

```{r}
s <- sample(D[, rowid], 12)
malAMLOrMDSTotal <- rowSums(D[, .(malAML, malMDS, malAMLOrMDS)], na.rm=TRUE)
D <- D[, malAMLOrMDSTotal := malAMLOrMDSTotal]
D <- D[is.na(malAML) & is.na(malMDS) & is.na(malAMLOrMDS), malAMLOrMDSTotal := NA]
D[s, .(rowid, malAML, malMDS, malAMLOrMDS, malAMLOrMDSTotal)]
```

Remove text from `nITT` column.

```{r}
D <- D[authorYear == "Misset (1996)" & arm == 2, nITT := "137"]
D <- D[authorYear == "Fumoleu (2003)" & arm == 1, nITT := "210"]
D <- D[authorYear == "Fumoleu (2003)" & arm == 2, nITT := "197"]
D <- D[authorYear == "Fumoleu (2003)" & arm == 3, nITT := "195"]
D <- D[, nITT := as.integer(nITT)]
```

Clean up the `regimen` and `dose` columns.

```{r}
D <- D[, regimen := gsub("\\s$", "", regimen)]
D <- D[, regimen := gsub("\\r\\n", " ", regimen)]
doseVar <- grep("dose", names(D), ignore.case = TRUE, value = TRUE)
for (i in 1:length(doseVar)) {
    D <- D[, doseVar[i] := gsub("\\r\\n", " ", get(doseVar[i]))]
}
D[, .N, c("regimen", doseVar), with = TRUE][order(regimen)]
write.table(D[, .N, c("regimen", doseVar), with = TRUE][order(regimen)],
            file = "regimens.md",
            sep = " | ", quote = FALSE,
            row.names = FALSE)
```

Calculate cumulative dose: total dose * number of courses.

```{r}
D <- D[, anthracyclineTotalDose := as.numeric(anthracyclineTotalDose)]
D <- D[authorYear == "Bergh (2000)" & regimen == "FEC\nTailored", anthracyclineTotalDose := 75]
D <- D[authorYear == "Henderson (2003)", anthracyclineTotalDose := 75]
D <- D[authorYear == "Colleoni (2009)" & regimen == "AC-CMF", anthracyclineTotalDose := 75]
D <- D[, cyclophosphamideDose := as.numeric(cyclophosphamideDose)]
D <- D[authorYear == "Bergh (2000)" & regimen == "FEC\nTailored", cyclophosphamideDose := 900]
D <- D[authorYear == "Henderson (2003)", anthracyclineTotalDose := 75]
D <- D[, taxaneCourses := as.numeric(taxaneCourses)]
D <- D[authorYear == "Shulman (2014)" & regimen == "T" & arm == 3, taxaneCourses := 4]
D <- D[authorYear == "Shulman (2014)" & regimen == "T" & arm == 4, taxaneCourses := 6]
D <- D[, fluoroucilTotalDose := as.numeric(fluoroucilTotalDose)]
D <- D[authorYear == "Bergh (2000)" & regimen == "FEC\nTailored", fluoroucilTotalDose := 600]
D <- D[authorYear == "Joensuu (2012)" & regimen == "TX-CEX", fluoroucilTotalDose := NA]
D <- D[,
       `:=` (anthracyclineCumulDose = anthracyclineTotalDose * anthracyclineCourses,
             cyclophosphamideCumulDose = cyclophosphamideDose * cyclophosphamideCourses,
             taxaneCumulDose = as.numeric(taxaneTotalDose) * taxaneCourses,
             fluoroucilCumulDose = fluoroucilTotalDose * fluoroucilCourses)]
D[, .(authorYear, regimen, anthracyclineCumulDose, anthracyclineTotalDose, anthracyclineCourses)]
D[, .(authorYear, regimen, cyclophosphamideCumulDose, cyclophosphamideDose, cyclophosphamideCourses)]
D[, .(authorYear, regimen, taxaneCumulDose, taxaneTotalDose, taxaneCourses)]
D[, .(authorYear, regimen, fluoroucilCumulDose, fluoroucilTotalDose, fluoroucilCourses)]
```


# Summarize

Define some functions.

```{r}
calcPct <- function (x, n) {
  prec <- "%.3g"
  sprintf(paste0(prec, "%%"),
          median(x / n, na.rm = TRUE) * 100)
}
calcRate <- function (x, n, y) {
  prec <- "%.3g"
  py <- 10000
  sprintf(paste(prec, "per %s p-y"),
          median(x / (n * (y / 12)), na.rm=TRUE) * py,
          py)
}
```

Summarize the regimens.

```{r}
D[,
  `:=`(isAnthra = !is.na(anthracyclineTotalDose),
        isCyclo = !is.na(cyclophosphamideDose),
        isTaxane = !is.na(taxaneTotalDose),
        isFluoro = !is.na(fluoroucilTotalDose))]
D1 <- melt(D,
           id.vars=c("id", "authorYear", "arm",
                     "isAnthra", "anthracyclineCumulDose",
                     "isCyclo", "cyclophosphamideCumulDose",
                     "isTaxane", "taxaneCumulDose",
                     "isFluoro", "fluoroucilCumulDose",
                     "nITT", "medianFU"),
           measure.vars=c("malAML", "malMDS", "malAMLOrMDSTotal", "malNonBreastSolid"),
           value.name="malN",
           variable.name="malType")
D1 <- D1[, malType := gsub("^mal", "", malType)]
D1 <- D1[, malType := factor(malType,
                             levels=c("AML", "MDS", "AMLOrMDSTotal", "NonBreastSolid"),
                             labels=c("AML", "MDS", "AML or MDS", "Non-Breast Solid"))]
D1 <- D1[, py := nITT * (medianFU / 12)]
D1 <- D1[, rate := malN / py * 10000]
summaryRegimens <- D1[,
                      .(totalN = sum(nITT, na.rm = TRUE),
                        totalPersonYears = round(sum(py, na.rm = TRUE)),
                        totalMalignancies = sum(malN, na.rm = TRUE),
                        medianPct = calcPct(malN, nITT),
                        medianRate = calcRate(malN, nITT, medianFU)),
                      .(isAnthra,
                        isCyclo,
                        isTaxane,
                        isFluoro,
                        malType)]
summaryRegimens <- summaryRegimens[order(-isAnthra, -isCyclo, -isTaxane, -isFluoro, malType)]
write.table(summaryRegimens,
            file = "summaryRegimens.md",
            sep = " | ", quote = FALSE,
            row.names = FALSE)
```


# Meta-regression

Estimate meta-regression models for log transformed incidence rate.

Define meta-regression functions.

```{r}
metareg <- function (D) {
  require(metafor)
  D <- D[!(is.na(x) | is.na(rate) | is.na(nITT) | is.na(malType))]
  M <- rma.glmm(xi = malN, ti = py, mods = xHighDose, measure="IRLN", data=D)
  pvalue <- M$pval[which(row.names(M$b) == "mods")]
  if (pvalue < 0.001) {
    pvalue <- "p < 0.001"
  } else {
    pvalue <- sprintf("p = %.03f", pvalue)
  }
  list(rma.glmm = M,
       pvalue = pvalue)
}
plotreg <- function (D, xlab, xbreaks, text, yhat) {
  require(ggplot2)
  require(tools)
  D <- D[!(is.na(x) | is.na(rate) | is.na(nITT) | is.na(malType))]
  D <- D[malType %in% c("AML or MDS", "Non-Breast Solid")]
  D <- D[, malType := droplevels(malType)]
  steps <- merge(D[, .(min = min(x), max = max(x)), .(malType, xHighDose)],
                 yhat,
                 by=c("malType", "xHighDose"))
  steps <- melt(steps, id.vars=c("malType", "xHighDose", "yhat"), measure.vars=c("min", "max"), value.name="x")
  G <- ggplot(D, aes(x=x, y=rate + 1/2, size=nITT / min(nITT, na.rm=TRUE), group=malType))
  G <- G + geom_point(alpha=1/2)
  G <- G + geom_smooth(method="lm", se=FALSE, linetype="dotted")
  G <- G + geom_step(data=steps, aes(x=x, y=yhat, group=malType), inherit.aes=FALSE, color="red")
  G <- G + geom_text(data=data.frame(x=Inf, y=1, label=text, malType=levels(D[, malType])),
                     aes(x, y, label=label, group=malType), inherit.aes=FALSE, hjust=1, color="red")
  G <- G + scale_x_log10(xlab, breaks=xbreaks)
  G <- G + scale_y_log10("Rate per 10,000 person-years")
  G <- G + facet_wrap(~ malType, nrow=2, ncol=1)
  G <- G + theme_bw()
  G <- G + theme(legend.position="none")
  filename <- gsub("\\s+", "", toTitleCase(xlab))
  ggsave(filename=sprintf("%s.png", filename))
  write.csv(D, file=sprintf("%s.csv", filename), row.names=FALSE)
  show(file.info(c(sprintf("%s.png", filename), sprintf("%s.csv", filename)))[c("size", "mtime")])
  G
}
invLogit <- function (x) {
  exp(x) / (1 + exp(x))
}
```

## Cyclophosphamide

Dichotomoize cumulative dose

* $\le 2400$
* $\gt 2400$

```{r}
library(metafor)
D2 <- D1[isCyclo == TRUE,
         .(id = factor(id),
           authorYear,
           isCyclo,
           x = cyclophosphamideCumulDose,
           xHighDose = cyclophosphamideCumulDose >= 2400,
           nITT,
           medianFU,
           malType,
           malN,
           py,
           rate)]
M1 <- metareg(D2[malType == "AML or MDS"      ])
M2 <- metareg(D2[malType == "Non-Breast Solid"])
pvalues <- c(M1$pvalue, M2$pvalue)
yhat1 <- unique(data.table(malType = "AML or MDS",
                           xHighDose = as.logical(M1$rma.glmm$X[, "mods"]),
                           yhat = invLogit(predict(M1$rma.glmm)[["pred"]]) * 10000))
yhat2 <- unique(data.table(malType = "Non-Breast Solid",
                           xHighDose = as.logical(M2$rma.glmm$X[, "mods"]),
                           yhat = invLogit(predict(M2$rma.glmm)[["pred"]]) * 10000))
yhat <- rbind(yhat1, yhat2)
plotreg(D2, "Cyclophosphamide cumulative dose", 1000 * c(0.5, 1, 2, 4, 8, 16), pvalues, yhat)
```


## Anthracycline

```{r, eval=FALSE}
library(metafor)
D2 <- D1[isAnthra == TRUE,
         .(id = factor(id),
           authorYear,
           isAnthra,
           x = anthracyclineCumulDose,
           nITT,
           medianFU,
           malType,
           malN,
           py,
           rate)]
M1 <- metareg(D2[malType == "AML or MDS"      ])
M2 <- metareg(D2[malType == "Non-Breast Solid"])
pvalues <- c(M1$pvalue, M2$pvalue)
plotreg(D2, "Anthracycline cumulative dose", 100 * c(1, 2, 4, 8), pvalues)
```


## Taxane

```{r, eval=FALSE}
library(metafor)
D2 <- D1[isTaxane == TRUE,
         .(id = factor(id),
           authorYear,
           isTaxane,
           x = taxaneCumulDose,
           nITT,
           medianFU,
           malType,
           malN,
           py,
           rate)]
M1 <- metareg(D2[malType == "AML or MDS"      ])
M2 <- metareg(D2[malType == "Non-Breast Solid"])
pvalues <- c(M1$pvalue, M2$pvalue)
plotreg(D2, "Taxane cumulative dose", 100 * c(1, 2, 4, 8), pvalues)
```


## Fluoroucil

```{r, eval=FALSE}
library(metafor)
D2 <- D1[isFluoro == TRUE,
         .(id = factor(id),
           authorYear,
           isFluoro,
           x = fluoroucilCumulDose,
           nITT,
           medianFU,
           malType,
           malN,
           py,
           rate)]
M1 <- metareg(D2[malType == "AML or MDS"      ])
M2 <- metareg(D2[malType == "Non-Breast Solid"])
pvalues <- c(M1$pvalue, M2$pvalue)
plotreg(D2, "Fluoroucil cumulative dose", 1000 * c(0.5, 1, 2, 4, 8, 16), pvalues)
```



# More stuff I'm not sure I need

```{r, eval=FALSE}
D3 <- D[isAnthra == TRUE & isCyclo == TRUE & isTaxane == FALSE & isFluoro == FALSE,
       .(isAnthra,
         isCyclo,
         isTaxane,
         isFluoro,
         malAML,
         nITT,
         medianFU,
         propAML = malAML / nITT,
         rateAML = malAML / medianFU)]
D3 <- D3[, medianPropAML := median(propAML, na.rm = TRUE)]
D3

# Hold the 1st sheet
D1 <- D

# Read the 2nd sheet
D <- read.xlsx(f, sheet = sheets[2], startRow = 1, colNames = TRUE)
D <- data.table(D)
oldnames <- names(D)
newnames <- c("id",
              "authorYear",
              "trial",
              "quality",
              "arm",
              "nRandomized",
              "nITT",
              "age",
              "menopausalStatus",
              "surgery",
              "tamoxifen",
              "tumorStage",
              "tumorSize",
              "nodalStatus",
              "hrStatus")
setnames(D, oldnames, newnames)


# Tidy data

# Input missing data for Bergh (2000); year is missing a digit
D[, authorYear := gsub("Bergh \\(200\\)", "Bergh \\(2000\\)", authorYear)]

# Recode values of "NR" and "-" to NA
D[, "menopausalStatus"] <- recode("menopausalStatus")
D[, "surgery"] <- recode("surgery")
D[, "tamoxifen"] <- recode("tamoxifen")
D[, "tumorStage"] <- recode("tumorStage")
D[, "tumorSize"] <- recode("tumorSize")
D[, "nodalStatus"] <- recode("nodalStatus")
D[, "hrStatus"] <- recode("hrStatus")


# Hold the 2nd sheet
D2 <- D


# Merge sheets
keyVar <- c("id", "authorYear", "arm")
D1 <- D1[, in1 := TRUE]
D2 <- D2[, in2 := TRUE]
setkeyv(D1, keyVar)
setkeyv(D2, keyVar)
D2 <- D2[,
         `:=` (trial = NULL,
               quality = NULL,
               nRandomized = NULL,
               nITT = NULL)]
D <- merge(D1, D2, all = TRUE)
D[is.na(in1) | is.na(in2), .(id, authorYear, arm, trial, in1, in2)]


# Save to RData
save(D, file = "SecMal.RData")
file.info("SecMal.RData")
load("SecMal.RData")
```
