# Meta-regression

Estimate meta-regression models for log transformed incidence rate.

Two outcomes are modeled separately.

1. AML or MDS secondary malignancies
1. Non-breast solid secondary malignancies

Cumulative doses are dichotomized into low and high doses.

* Cyclophosphamide
    * $\lt 2400 \frac{\text{mg}}{\text{m}^2}$
    * $\ge 2400 \frac{\text{mg}}{\text{m}^2}$
* Taxane
    * $\lt 500 \frac{\text{mg}}{\text{m}^2}$
    * $\ge 500 \frac{\text{mg}}{\text{m}^2}$

```{r}
xscale <- 1e3
D2 <- D1[,
         .(id = factor(id),
           authorYear,
           arm,
           regimen,
           isCyclo,
           xCyc = cyclophosphamideCumulDose / xscale,  # scale units
           isCycHighDose = cyclophosphamideCumulDose >= 2400,
           xTax = taxaneCumulDose / 1e2,  # scale units
           isTaxDoseHigh = ifelse(taxaneCumulDose >= 500 & isTaxane == TRUE, TRUE, FALSE),
           isTaxDoseLow = ifelse(taxaneCumulDose < 500 & isTaxane == TRUE, TRUE, FALSE),
           isDocetaxel,
           isPaclitaxel,
           isAnthra,
           isTaxane,
           isNoTaxane = !(isTaxane),
           isFluoro,
           nITT,
           medianFU,
           malType,
           malN,
           py,
           rate)]
D2 <- D2[isTaxane == FALSE, isTaxDoseHigh := FALSE]
D2 <- drop_na(D2, xCyc, isCycHighDose, isDocetaxel, isPaclitaxel)
```

The effect of cyclophosphamide could be due to high dose taxanes in the regimen.

> From: Joy Melnikow [mailto:jamelnikow@ucdavis.edu]   
> Sent: Monday, August 01, 2016 2:47 PM  
> To: Benjamin Chan <chanb@ohsu.edu>  
> Subject: RE: secondary malignancies  
> 
> The one I'm asking about is cyclophosphamide with high dose taxane vs
> cyclophosphamide alone.  I think if it looks like taxane augments the known
> risk of cyclophosphamide this would be an important addition to the
> literature.

The following models are estimated.

1. $$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \beta_3 I_{\text{docetaxel}, i} + \sigma_j$$
1. $$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \sigma_j$$
1. $$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \beta_4 I_{\text{no taxane}, i} + \sigma_j$$
1. $$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \gamma x_{1, i} I_{\text{paclitaxel}, i} + \sigma_j$$

where

* $x_{1, i}$ is the cumulative cyclophosphamide dose for the $i$-th treatment arm
* $I_{\text{paclitaxel}, i}$ is an indicator for whether the $i$-th treatment arm included paclitaxel is the regimen
* $I_{\text{docetaxel}, i}$ is an indicator for whether the $i$-th treatment arm included docetaxel is the regimen
* $I_{\text{no taxane}, i}$ is an indicator for whether the $i$-th treatment arm did not include any taxane is the regimen
* $\sigma_j$ is a random effect for the $j$-th study.

Models were estimated using the `rma.mv()` function from the metafor` package for R.

```{r}
library(metafor)
citation("metafor")
```

Define meta-regression functions.


```{r}
pvalToChar <- function (p) {
  if (p < 0.001) {
    pvalue <- "p < 0.001"
  } else {
    pvalue <- sprintf("p = %.02g", p)
  }
  pvalue
}
metareg <- function (D) {
  require(metafor)
  xData <- unique(D[, .(xCyc, isCycHighDose, isDocetaxel, isPaclitaxel, malType)])
  D <- escalc("IRLN", xi = malN, ti = py, data = D)
  randomEffect <- list(~ 1 | id)
  M1 <- rma.mv(yi ~ xCyc + isPaclitaxel + isDocetaxel,
               vi,
               random = randomEffect,
               data = D)
  M2 <- rma.mv(yi ~ xCyc + isPaclitaxel,
               vi,
               random = randomEffect,
               data = D)
  M3 <- rma.mv(yi ~ xCyc + isPaclitaxel + isNoTaxane,
               vi,
               random = randomEffect,
               data = D)
  M4 <- rma.mv(yi ~ xCyc + isPaclitaxel + xCyc * isPaclitaxel,
               vi,
               random = randomEffect,
               data = D)
  list(M1 = M1, M2 = M2, M3 = M3, M4 = M4)
}
plotreg <- function (M, D, title) {
  require(ggplot2)
  require(svglite)
  require(RColorBrewer)
  require(data.table)
  D <- data.table(D)
  X <- data.table(M$X)
  X <- unique(X)
  varnames <- names(X)[2:ncol(X)]
  X1 <- data.table(x1 = seq(min(X[isPaclitaxelTRUE == 0 & isDocetaxelTRUE == 0, xCyc]),
                            max(X[isPaclitaxelTRUE == 0 & isDocetaxelTRUE == 0, xCyc]),
                            length.out = 100),
                   x2 = rep(0, 100),
                   x3 = rep(0, 100))
  X2 <- data.table(x1 = seq(min(X[isPaclitaxelTRUE == 1 & isDocetaxelTRUE == 0, xCyc]),
                            max(X[isPaclitaxelTRUE == 1 & isDocetaxelTRUE == 0, xCyc]),
                            length.out = 100),
                   x2 = rep(1, 100),
                   x3 = rep(0, 100))
  X3 <- data.table(x1 = seq(min(X[isPaclitaxelTRUE == 0 & isDocetaxelTRUE == 1, xCyc]),
                            max(X[isPaclitaxelTRUE == 0 & isDocetaxelTRUE == 1, xCyc]),
                            length.out = 100),
                   x2 = rep(0, 100),
                   x3 = rep(1, 100))
  X <- rbindlist(list(X1, X2, X3))
  names(X) <- varnames
  X$xCyc <- round(X$xCyc, digits = 1)
  X <- unique(X)
  yhat <- data.table(malType = mal,
                     X,
                     # variable = "pred",
                     pred = predict(M, as.matrix(X), transf = exp)[["pred"]] * scale)
  ci.lb <- data.table(malType = mal,
                      X,
                      # variable = "ci.lb",
                      ci.lb = predict(M, as.matrix(X), transf = exp)[["ci.lb"]] * scale)
  ci.ub <- data.table(malType = mal,
                      X,
                      # variable = "ci.ub",
                      ci.ub = predict(M, as.matrix(X), transf = exp)[["ci.ub"]] * scale)
  keyvar <- c("malType", "xCyc", "isPaclitaxelTRUE", "isDocetaxelTRUE")
  setkeyv(yhat, keyvar)
  setkeyv(ci.lb, keyvar)
  setkeyv(ci.ub, keyvar)
  yhat <- merge(merge(yhat, ci.lb), ci.ub)
  pvalues <- c(M$pval[grep("TRUE$", row.names(M$b))])
  names(pvalues) <- row.names(M$b)[grep("TRUE$", row.names(M$b))]
  pal <- brewer.pal(ncol(X), name = "Set1")
  D <- D[isPaclitaxel == FALSE & isDocetaxel == FALSE, color := "A"]
  D <- D[isPaclitaxel == TRUE  & isDocetaxel == FALSE, color := "B"]
  D <- D[isPaclitaxel == FALSE & isDocetaxel == TRUE , color := "C"]
  D <- D[, color := factor(color, labels = c("A", "B", "C"))]
  setorder(D, color)
  G <- ggplot(D, aes(x = xCyc * xscale,
                     y = rate + 1/2,
                     size = nITT / min(nITT, na.rm = TRUE),
                     color = color))
  G <- G + geom_point(alpha = 1/2,
                      position = "jitter")
  G <- G + geom_line(data = yhat[isDocetaxelTRUE == 0 & isPaclitaxelTRUE == 0, ],
                     aes(x = xCyc * xscale, y = pred),
                     inherit.aes = FALSE,
                     color = pal[1])
  G <- G + geom_line(data = yhat[isDocetaxelTRUE == 0 & isPaclitaxelTRUE == 1, ],
                     aes(x = xCyc * xscale, y = pred),
                     inherit.aes = FALSE,
                     color = pal[2])
  G <- G + geom_line(data = yhat[isDocetaxelTRUE == 1 & isPaclitaxelTRUE == 0, ],
                     aes(x = xCyc * xscale, y = pred),
                     inherit.aes = FALSE,
                     color = pal[3])
  G <- G + geom_ribbon(data = yhat[isDocetaxelTRUE == 0 & isPaclitaxelTRUE == 0, ],
                       aes(x = xCyc * xscale, ymax = ci.ub, ymin = ci.lb),
                       inherit.aes = FALSE,
                       fill = pal[1],
                       alpha = 1/8)
  G <- G + geom_ribbon(data = yhat[isDocetaxelTRUE == 0 & isPaclitaxelTRUE == 1, ],
                       aes(x = xCyc * xscale, ymax = ci.ub, ymin = ci.lb),
                       inherit.aes = FALSE,
                       fill = pal[2],
                       alpha = 1/8)
  G <- G + geom_ribbon(data = yhat[isDocetaxelTRUE == 1 & isPaclitaxelTRUE == 0, ],
                       aes(x = xCyc * xscale, ymax = ci.ub, ymin = ci.lb),
                       inherit.aes = FALSE,
                       fill = pal[3],
                       alpha = 1/8)
  G <- G + scale_x_log10("Cyclophosphamide cumulative dose",
                         breaks = 1e3 * (2 ^ seq(-1, 6)))
  G <- G + scale_y_log10(sprintf("Rate per %s person-years",
                                 format(scale, big.mark = ",")),
                         breaks = 2 ^ seq(-1, 6))
  G <- G + scale_color_manual("Taxane",
                              values = pal,
                              labels = c("None (reference)",
                                         sprintf("Paclitaxel (%s)",
                                                 pvalToChar(pvalues[grep("Pacli", names(pvalues))])),
                                         sprintf("Docetaxel (%s)",
                                                 pvalToChar(pvalues[grep("Doce", names(pvalues))]))))
  G <- G + scale_size_continuous(guide = FALSE)
  G <- G + labs(title = title)
  G <- G + theme_bw()
  filename <- sprintf("%s_Cyclophosphamide_byTaxane",
                      gsub("(\\s)|(-)", "", title))
  ggsave(filename = sprintf("%s.png", filename), width = 9)
  ggsave(filename = sprintf("%s.svg", filename), width = 9)
  yhat$xCyc <- yhat$xCyc * xscale
  write.csv(D, file = sprintf("%s.csv", filename), row.names = FALSE, quote = FALSE)
  write.csv(yhat, file = sprintf("%s_Pred.csv", filename), row.names = FALSE, quote = FALSE)
  show(file.info(grep(paste0(filename, "(_Pred)*\\."), list.files(), value = TRUE))[c("size", "mtime")])
  G
}
```

## AML/MDS

```{r}
mal <- "AML or MDS"
D3 <- drop_na(D2[malType == mal], rate, nITT, malType)
M <- metareg(D3)
```

### Findings

* Crude statistics
  * Outcome: `r mal`
  * Total person-years of follow-up: `r D3[, format(round(sum(py)), big.mark = ",")]`
  * Total number of persons (ITT): `r D3[, format(sum(nITT), big.mark = ",")]`
  * Total number of outcomes: `r D3[, format(sum(malN), big.mark = ",")]`
  * Number of treatment arms: `r nrow(D3)`
  * Number of studies: `r length(unique(D3$id))`
  * Crude incidence rate: `r D3[, sprintf("%.03g per %s person-years", sum(malN) / sum(py) * scale, format(scale, big.mark = ","))]`
* `r mal` rate had a dose response relationship with cumulative cyclophosphamide dose (M1)
  * `r mal` rate increased `r sprintf("%.03g", exp(M$M1$b[which(row.names(M$M1$b) == "xCyc")]))` times (95% CI: `r sprintf("%.03g", exp(M$M1$ci.lb[which(row.names(M$M1$b) == "xCyc")]))`, `r sprintf("%.03g", exp(M$M1$ci.ub[which(row.names(M$M1$b) == "xCyc")]))`; p = `r sprintf("%.02g", M$M1$pval[which(row.names(M$M1$b) == "xCyc")])`) for each `r xscale` $\text{mg} / \text{m}^2$
* Paclitaxel confounded the cyclophosphamide dose response
  * Compared to no taxane (M1), paclitaxel increased `r mal` rate by `r sprintf("%.03g", exp(M$M1$b[which(row.names(M$M1$b) == "isPaclitaxelTRUE")]))` times (95% CI: `r sprintf("%.03g", exp(M$M1$ci.lb[which(row.names(M$M1$b) == "isPaclitaxelTRUE")]))`, `r sprintf("%.03g", exp(M$M1$ci.ub[which(row.names(M$M1$b) == "isPaclitaxelTRUE")]))`; p = `r sprintf("%.02g", M$M1$pval[which(row.names(M$M1$b) == "isPaclitaxelTRUE")])`)
  * Compared to no taxane or docetaxel (M2), paclitaxel increased `r mal` rate by `r sprintf("%.03g", exp(M$M2$b[which(row.names(M$M2$b) == "isPaclitaxelTRUE")]))` times (95% CI: `r sprintf("%.03g", exp(M$M2$ci.lb[which(row.names(M$M2$b) == "isPaclitaxelTRUE")]))`, `r sprintf("%.03g", exp(M$M2$ci.ub[which(row.names(M$M2$b) == "isPaclitaxelTRUE")]))`; p = `r sprintf("%.02g", M$M2$pval[which(row.names(M$M2$b) == "isPaclitaxelTRUE")])`)
  * Compared to docetaxel (M3), paclitaxel increased `r mal` rate by `r sprintf("%.03g", exp(M$M3$b[which(row.names(M$M3$b) == "isPaclitaxelTRUE")]))` times (95% CI: `r sprintf("%.03g", exp(M$M3$ci.lb[which(row.names(M$M3$b) == "isPaclitaxelTRUE")]))`, `r sprintf("%.03g", exp(M$M3$ci.ub[which(row.names(M$M3$b) == "isPaclitaxelTRUE")]))`; p = `r sprintf("%.02g", M$M3$pval[which(row.names(M$M3$b) == "isPaclitaxelTRUE")])`)
* Paclitaxel did not modify the dose response effect of cyclophosphamide (M4)
  * Interaction term estimate was `r sprintf("%.03g", M$M4$b[which(row.names(M$M4$b) == "xCyc:isPaclitaxelTRUE")])` (95% CI: `r sprintf("%.03g", M$M4$ci.lb[which(row.names(M$M4$b) == "xCyc:isPaclitaxelTRUE")])`, `r sprintf("%.03g", M$M4$ci.ub[which(row.names(M$M4$b) == "xCyc:isPaclitaxelTRUE")])`; p = `r sprintf("%.02g", M$M4$pval[which(row.names(M$M4$b) == "xCyc:isPaclitaxelTRUE")])`)

### M1

* Cyclophosphamide: dose response
* Paclitaxel and docetaxel: confounders

$$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \beta_3 I_{\text{docetaxel}, i} + \sigma_j$$

```{r, fig.width = 9}
M$M1
plotreg(M$M1, D3, mal)
```

### M2

* Cyclophosphamide: dose response
* Paclitaxel: confounder

$$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \sigma_j$$

```{r}
M$M2
```

### M3

* Cyclophosphamide: dose response
* Paclitaxel and no taxane: confounders

$$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \beta_4 I_{\text{no taxane}, i} + \sigma_j$$

```{r}
M$M3
```

### M4

* Cyclophosphamide: dose response
* Paclitaxel: effect modifier

$$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \gamma x_{1, i} I_{\text{paclitaxel}, i} + \sigma_j$$

```{r}
M$M4
```


## Non-Breast Solid

```{r}
mal <- "Non-Breast Solid"
D3 <- drop_na(D2[malType == mal], rate, nITT, malType)
M <- metareg(D3)
```

### Findings

* Crude statistics
  * Outcome: `r mal`
  * Total person-years of follow-up: `r D3[, format(round(sum(py)), big.mark = ",")]`
  * Total number of persons (ITT): `r D3[, format(sum(nITT), big.mark = ",")]`
  * Total number of outcomes: `r D3[, format(sum(malN), big.mark = ",")]`
  * Number of treatment arms: `r nrow(D3)`
  * Number of studies: `r length(unique(D3$id))`
  * Crude incidence rate: `r D3[, sprintf("%.03g per %s person-years", sum(malN) / sum(py) * scale, format(scale, big.mark = ","))]`
* `r mal` rate did not have a dose response relationship with cumulative cyclophosphamide dose (M1)
  * `r mal` rate increased `r sprintf("%.03g", exp(M$M1$b[which(row.names(M$M1$b) == "xCyc")]))` times (95% CI: `r sprintf("%.03g", exp(M$M1$ci.lb[which(row.names(M$M1$b) == "xCyc")]))`, `r sprintf("%.03g", exp(M$M1$ci.ub[which(row.names(M$M1$b) == "xCyc")]))`; p = `r sprintf("%.02g", M$M1$pval[which(row.names(M$M1$b) == "xCyc")])`) for each `r xscale` $\text{mg} / \text{m}^2$
* Paclitaxel did not confound the cyclophosphamide dose response
  * Compared to no taxane (M1), paclitaxel increased `r mal` rate by `r sprintf("%.03g", exp(M$M1$b[which(row.names(M$M1$b) == "isPaclitaxelTRUE")]))` times (95% CI: `r sprintf("%.03g", exp(M$M1$ci.lb[which(row.names(M$M1$b) == "isPaclitaxelTRUE")]))`, `r sprintf("%.03g", exp(M$M1$ci.ub[which(row.names(M$M1$b) == "isPaclitaxelTRUE")]))`; p = `r sprintf("%.02g", M$M1$pval[which(row.names(M$M1$b) == "isPaclitaxelTRUE")])`)
  * Compared to no taxane or docetaxel (M2), paclitaxel increased `r mal` rate by `r sprintf("%.03g", exp(M$M2$b[which(row.names(M$M2$b) == "isPaclitaxelTRUE")]))` times (95% CI: `r sprintf("%.03g", exp(M$M2$ci.lb[which(row.names(M$M2$b) == "isPaclitaxelTRUE")]))`, `r sprintf("%.03g", exp(M$M2$ci.ub[which(row.names(M$M2$b) == "isPaclitaxelTRUE")]))`; p = `r sprintf("%.02g", M$M2$pval[which(row.names(M$M2$b) == "isPaclitaxelTRUE")])`)
  * Compared to docetaxel (M3), paclitaxel increased `r mal` rate by `r sprintf("%.03g", exp(M$M3$b[which(row.names(M$M3$b) == "isPaclitaxelTRUE")]))` times (95% CI: `r sprintf("%.03g", exp(M$M3$ci.lb[which(row.names(M$M3$b) == "isPaclitaxelTRUE")]))`, `r sprintf("%.03g", exp(M$M3$ci.ub[which(row.names(M$M3$b) == "isPaclitaxelTRUE")]))`; p = `r sprintf("%.02g", M$M3$pval[which(row.names(M$M3$b) == "isPaclitaxelTRUE")])`)
* Paclitaxel did not modify the dose response effect of cyclophosphamide (M4)
  * Interaction term estimate was `r sprintf("%.03g", M$M4$b[which(row.names(M$M4$b) == "xCyc:isPaclitaxelTRUE")])` (95% CI: `r sprintf("%.03g", M$M4$ci.lb[which(row.names(M$M4$b) == "xCyc:isPaclitaxelTRUE")])`, `r sprintf("%.03g", M$M4$ci.ub[which(row.names(M$M4$b) == "xCyc:isPaclitaxelTRUE")])`; p = `r sprintf("%.02g", M$M4$pval[which(row.names(M$M4$b) == "xCyc:isPaclitaxelTRUE")])`)

### M1

* Cyclophosphamide: dose response
* Paclitaxel and docetaxel: confounders

$$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \beta_3 I_{\text{docetaxel}, i} + \sigma_j$$

```{r, fig.width = 9}
M$M1
plotreg(M$M1, D3, mal)
```

### M2

* Cyclophosphamide: dose response
* Paclitaxel: confounder

$$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \sigma_j$$

```{r}
M$M2
```

### M3

* Cyclophosphamide: dose response
* Paclitaxel and no taxane: confounders

$$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \beta_4 I_{\text{no taxane}, i} + \sigma_j$$

```{r}
M$M3
```

### M4

* Cyclophosphamide: dose response
* Paclitaxel: effect modifier

$$\log \left( \frac{y_i}{t_i} \right) = \beta_0 + \beta_1 x_{1, i} + \beta_2 I_{\text{paclitaxel}, i} + \gamma x_{1, i} I_{\text{paclitaxel}, i} + \sigma_j$$

```{r}
M$M4
```
