# Meta-regression

Estimate meta-regression models for log transformed incidence rate.
Model is

$$\frac{y_i}{t_i} = \beta x_i + \sigma_\text{study}$$

Or

$$\frac{y_i}{t_i} = \beta I_{\text{high dose}, i} + \sigma_\text{study}$$

Models were estimated using the `rma.mv()` function from the metafor` package for R.


```{r}
library(metafor)
citation("metafor")
```

Define meta-regression functions.


```{r}
pvalToChar <- function (p) {
  if (p < 0.001) {
    pvalue <- "p < 0.001"
  } else {
    pvalue <- sprintf("p = %.03f", p)
  }
  pvalue
}
metareg <- function (D) {
  require(metafor)
  D <- D[!(is.na(x) | is.na(rate) | is.na(nITT) | is.na(malType))]
  xData <- unique(D[, .(drug, x, xHighDose, malType)])
  D <- escalc("IRLN", xi=malN, ti=py, data=D)
  randomEffect <- list(~ 1 | id)
  MLin <- rma.mv(yi ~ x, vi, random=randomEffect, data=D)
  MBin <- rma.mv(yi ~ xHighDose, vi, random=randomEffect, data=D)
  pvalueLin <- MLin$pval[which(row.names(MLin$b) == "x")]
  pvalueBin <- MBin$pval[which(row.names(MBin$b) == "xHighDoseTRUE")]
  predLin <- predict(MLin, xData[, x], transf = exp)[["pred"]] * scale
  confLowerLin <- predict(MLin, xData[, x], transf = exp)[["cr.lb"]] * scale
  confUpperLin <- predict(MLin, xData[, x], transf = exp)[["cr.ub"]] * scale
  predBin <- predict(MBin, xData[, xHighDose], transf = exp)[["pred"]] * scale
  confLowerBin <- predict(MBin, xData[, xHighDose], transf = exp)[["cr.lb"]] * scale
  confUpperBin <- predict(MBin, xData[, xHighDose], transf = exp)[["cr.ub"]] * scale
  pred <- data.table(xData, predLin, confLowerLin, confUpperLin, predBin, confLowerBin, confUpperBin, scale)
  setorder(pred, x)
  list(rmaLin = MLin,
       rmaBin = MBin,
       pvalueLin = pvalToChar(pvalueLin),
       pvalueBin = pvalToChar(pvalueBin),
       pred = pred)
}
plotreg <- function (M, D, title, xlab, xbreaks, xscale) {
  require(ggplot2)
  require(RColorBrewer)
  require(tools)
  pvalues <- c(M$pvalueLin, M$pvalueBin)
  x <- seq(min(M$rmaLin$X[, "x"]), max(M$rmaLin$X[, "x"]), length.out=100)
  yhat1 <- unique(data.table(malType = mal,
                             x,
                             yhat = predict(M$rmaLin, x, transf = exp)[["pred"]] * scale))
  xHighDose <- as.logical(M$rmaBin$X[, "xHighDoseTRUE"])
  yhat2 <- unique(data.table(malType = mal,
                             xHighDose,
                             yhat = predict(M$rmaBin, xHighDose, transf = exp)[["pred"]] * scale))
  D <- D[!(is.na(x) | is.na(rate) | is.na(nITT) | is.na(malType))]
  D <- D[, malType := droplevels(malType)]
  steps <- merge(D[, .(min = min(x), max = max(x)), .(malType, xHighDose)],
                 yhat2,
                 by=c("malType", "xHighDose"))
  steps <- melt(steps, id.vars=c("malType", "xHighDose", "yhat"), measure.vars=c("min", "max"), value.name="x")
  annoLin <- data.frame(x=Inf, y=1.2, label=pvalues[c(1)], malType=levels(D[, malType]))
  annoBin <- data.frame(x=Inf, y=1.4, label=pvalues[c(2)], malType=levels(D[, malType]))
  pal <- brewer.pal(4, name="RdBu")
  G <- ggplot(D, aes(x=x * xscale, y=rate + 1/2, size=nITT / min(nITT, na.rm=TRUE)))
  G <- G + geom_point(alpha=1/3, position="jitter")
  G <- G + geom_line(data=yhat1, aes(x=x * xscale, y=yhat), inherit.aes=FALSE, color=pal[4])
  G <- G + geom_step(data=steps, aes(x=x * xscale, y=yhat), inherit.aes=FALSE, color=pal[1])
  G <- G + geom_text(data=annoLin,
                     aes(x, y, label=label), inherit.aes=FALSE, hjust=1, color=pal[4])
  G <- G + geom_text(data=annoBin,
                     aes(x, y, label=label), inherit.aes=FALSE, hjust=1, color=pal[1])
  G <- G + scale_x_log10(xlab, breaks=xbreaks)
  G <- G + scale_y_log10(sprintf("Rate per %s person-years", format(scale, big.mark=",")))
  G <- G + labs(title=title)
  G <- G + theme_bw()
  G <- G + theme(legend.position="none")
  filename <- gsub("\\s+",
                   "",
                   toTitleCase(paste(gsub("[[:punct:]]", "", title),
                                     gsub("cumulative dose", "", xlab),
                                     sep="_")))
  ggsave(filename=sprintf("%s.png", filename))
  write.csv(D, file=sprintf("%s.csv", filename), row.names=FALSE, quote=FALSE)
  M$pred[, x := x * xscale]
  write.csv(M$pred, file=sprintf("%s_Pred.csv", filename), row.names=FALSE, quote=FALSE)
  show(file.info(grep(paste0(filename, "(_Pred)*\\."), list.files(), value = TRUE))[c("size", "mtime")])
  G
}
```

Dichotomize cumulative doses

* Cyclophosphamide
    * $\lt 2400$
    * $\ge 2400$
* Taxane
    * $\lt 500$
    * $\ge 500$


```{r}
D2 <- D1[,
         .(id = factor(id),
           authorYear,
           isCyclo,
           xCyc = cyclophosphamideCumulDose / 1e3,
           xCycHighDose = cyclophosphamideCumulDose >= 2400,
           xTax = taxaneCumulDose / 1e2,
           xTaxHighDose = taxaneCumulDose >= 500,
           isAnthra,
           isTaxane,
           isFluoro,
           nITT,
           medianFU,
           malType,
           malN,
           py,
           rate)]
```

## AML/MDS

* Fit the models
* Show model summaries
* Plot the predicted values


```{r}
mal <- "AML or MDS"
D3 <- D2[malType == mal]
```


### Cyclophosphamide

```{r}
D3 <- D3[, `:=` (drug = "Cyclophosphamide", x = xCyc, xHighDose = xCycHighDose)]
M <- metareg(D3)
M
plotreg(M,
        D3,
        mal,
        "Cyclophosphamide cumulative dose",
        1e3 * c(0.5, 1, 2, 4, 8, 16),
        1e3)
```

### Taxane

```{r}
D3 <- D3[, `:=` (drug = "Taxane", x = xTax, xHighDose = xTaxHighDose)]
M <- metareg(D3)
M
plotreg(M,
        D3,
        mal,
        "Taxane cumulative dose",
        1e2 * c(0.5, 1, 2, 4, 8, 16),
        1e2)
```

### Cyclophosphamide, with and without High Dose Taxane

The effect of cyclophosphamide could be due to high dose taxanes in the regimen.

> From: Joy Melnikow [mailto:jamelnikow@ucdavis.edu]   
> Sent: Monday, August 01, 2016 2:47 PM  
> To: Benjamin Chan <chanb@ohsu.edu>  
> Subject: RE: secondary malignancies  
> 
> The one I'm asking about is cyclophosphamide with high dose taxane vs
> cyclophosphamide alone.  I think if it looks like taxane augments the known
> risk of cyclophosphamide this would be an important addition to the
> literature.


Fit the meta-regression model

$$\frac{y_i}{t_i} = \beta_0 + \beta_1 x_{\text{cyclophosphamide dose}, i} + \beta_2 I_{\text{high dose taxane}, i} + \sigma_\text{study}$$

where $I_{\text{high dose taxane}, i}$ is an 0/1 indicator for whether study $i$ has high dose taxane in the regimen.
The interpretation of $\beta_1$ is the effect of cyclophosphamide dose adjusted for high dose taxane in the regimen.

```{r}
mal <- "AML or MDS"
D3 <- D2[malType == mal]
D3 <- D3[, drug := "Cyclophosphamide"]
D3 <- D3[isCyclo == FALSE, xCycHighDose := FALSE]
D3 <- D3[, isTaxaneHighDose := FALSE]
D3 <- D3[xTaxHighDose == TRUE, isTaxaneHighDose := TRUE]
D3 <- D3[!(is.na(xCyc) | is.na(xCycHighDose) | is.na(rate) | is.na(nITT) | is.na(malType))]
xData <- unique(D3[, .(drug, xCyc, xCycHighDose, isTaxaneHighDose, malType)])
D3 <- escalc("IRLN", xi=malN, ti=py, data=D3)
randomEffect <- list(~ 1 | id)
MLin <- rma.mv(yi ~ xCyc + isTaxaneHighDose,
               vi,
               random=randomEffect,
               data=D3)
MLinInt <- rma.mv(yi ~ xCyc + isTaxaneHighDose + xCyc * isTaxaneHighDose,
                  vi,
                  random=randomEffect,
                  data=D3)
MBin <- rma.mv(yi ~ xCycHighDose + isTaxaneHighDose,
               vi,
               random=randomEffect,
               data=D3)
M <- list(rmaLin = MLin,
          rmaLinInt = MLinInt,
          rmaBin = MBin)
M
```

Plot predicted values.

```{r}
X <- data.frame(M$rmaLin$X)
X <- unique(X)
varnames <- names(X)[2:ncol(X)]
X0 <- data.frame(x1 = seq(min(X[X[, 3] == 0, "xCyc"]), max(X[X[, 3] == 0, "xCyc"]),
                          length.out = 100),
                 x2 = rep(0, 100))
X1 <- data.frame(x1 = seq(min(X[X[, 3] == 1, "xCyc"]), max(X[X[, 3] == 1, "xCyc"]),
                          length.out = 100),
                 x2 = rep(1, 100))
X <- rbind(X0, X1)
names(X) <- varnames
X$xCyc <- round(X$xCyc, digits = 1)
X <- unique(X)
yhat <- data.frame(malType = mal,
                   X,
                   yhat = predict(M$rmaLin, as.matrix(X), transf = exp)[["pred"]] * scale)
xscale <- 1e3
pvalues <- c(M$rmaLin$pval[which(row.names(M$rmaLin$b) == "isTaxaneHighDoseTRUE")])
pal <- brewer.pal(4, name = "PuOr")
G <- ggplot(D3, aes(x = xCyc * xscale,
                    y = rate + 1/2,
                    size = nITT / min(nITT, na.rm=TRUE),
                    color = isTaxaneHighDose))
G <- G + geom_point(alpha = 1/2,
                    position = "jitter")
G <- G + geom_line(data = yhat[yhat$isTaxaneHighDose == FALSE, ],
                   aes(x = xCyc * xscale, y = yhat),
                   inherit.aes = FALSE,
                   color = pal[1])
G <- G + geom_line(data = yhat[yhat$isTaxaneHighDose == TRUE, ],
                   aes(x = xCyc * xscale, y = yhat),
                   inherit.aes = FALSE,
                   color = pal[4])
G <- G + scale_x_log10("Cyclophosphamide cumulative dose",
                       breaks = 1e3 * c(0.5, 1, 2, 4, 8, 16))
G <- G + scale_y_log10(sprintf("Rate per %s person-years",
                               format(scale, big.mark = ",")))
G <- G + scale_color_manual(sprintf("Taxane, %s", pvalToChar(pvalues[c(1)])),
                            values = c(pal[1], pal[4]),
                            labels = c("None or low dose", "High dose"))
G <- G + scale_size_continuous(guide = FALSE)
G <- G + labs(title = mal)
G <- G + theme_bw()
filename <- "AMLorMDS_Cyclophosphamide_byHighDoseTaxane"
ggsave(filename = sprintf("%s.png", filename), width = 9)
yhat$xCyc <- yhat$xCyc * xscale
write.csv(yhat, file = sprintf("%s_Pred.csv", filename), row.names = FALSE, quote = FALSE)
show(file.info(grep(paste0(filename, "(_Pred)*\\."), list.files(), value = TRUE))[c("size", "mtime")])
G
```


## Non-Breast Solid

* Fit the models
* Show model summaries
* Plot the predicted values


```{r}
mal <- "Non-Breast Solid"
D3 <- D2[malType == mal]
```

### Cyclophosphamide

```{r}
D3 <- D3[, `:=` (drug = "Cyclophosphamide", x = xCyc, xHighDose = xCycHighDose)]
M <- metareg(D3)
M
plotreg(M,
        D3,
        mal,
        "Cyclophosphamide cumulative dose",
        1e3 * c(0.5, 1, 2, 4, 8, 16),
        1e3)
```

### Taxane

```{r}
D3 <- D3[, `:=` (drug = "Taxane", x = xTax, xHighDose = xTaxHighDose)]
M <- metareg(D3)
M
plotreg(M,
        D3,
        mal,
        "Taxane cumulative dose",
        1e2 * c(0.5, 1, 2, 4, 8, 16),
        1e2)
```
