# Load data

Load the data from [GitHub](https://github.com/benjamin-chan/SecMalAfterBreastCaACT).


```{r}
repo <- "https://github.com/benjamin-chan/SecMalAfterBreastCaACT"
url <- paste0(repo, "/raw/26115dab1eb2b075d40f545514a03c3b30cde6b4/SecondaryMalignancies_Data_BG_3_4.xlsx")
f <- tempfile()
download.file(url, f, mode="wb")
file.info(f)[c("size", "mtime")]
sheets <- getSheetNames(f)
# sheets
D <- read.xlsx(f, sheet = sheets[1], startRow = 3, colNames = FALSE)
D <- data.table(D)
oldnames <- names(D)
newnames <- c("id",
              "authorYear",
              "trial",
              "quality",
              "arm",
              "nRandomized",
              "nITT",
              # "pctCompletingTx",
              "medianFU",
              "regimen",
              "anthracyclineType",
              "anthracyclineTotalDose",
              "anthracyclineDuration",
              "anthracyclineCourses",
              "cyclophosphamideDose",
              "cyclophosphamideDuration",
              "cyclophosphamideCourses",
              "taxaneType",
              "taxaneTotalDose",
              "taxaneDuration",
              "taxaneCourses",
              "fluoroucilTotalDose",
              "fluoroucilDuration",
              "fluoroucilCourses",
              "otherTxDetails",
              "malAML",
              "malMDS",
              "malAMLOrMDS",
              "malNonBreastSolid",
              "malNonBreastSolidType",
              "malOtherBloodCancers",
              "malSMRelatedDeaths",
              "malSecondPrimary",
              "NOTES")
setnames(D, oldnames, newnames[1:33])
```

# Tidy data

Create a `rowid` variable.
This will be handy later.


```{r}
D <- D[, rowid := .I]
```

Fix some minor data entry inconsistencies.


```{r}
D <- D[grep("Nitz", authorYear), authorYear := gsub("2014$", "2014)", authorYear)]
D <- D[grep("Wolmark", authorYear), authorYear := gsub(", 2001$", " (2001)", authorYear)]
```

Input missing data for Bergh (2000); row was split.


```{r}
col <- grep("^cyclophosphamide", names(D), invert = TRUE)
D[11, col] <- D[10, col, with = FALSE]
# D[11, ]
# D[10, ]
```

Fix trial column for Romond (2005).


```{r}
D <- D[grep("Romond", authorYear), trial := gsub("&N", "& N", trial)]
```

Make `medianFU` numeric.


```{r}
D <- D[grep("-", medianFU, invert = TRUE), medianFUNum := round(as.numeric(medianFU), digits = 2)]
x <- D[grep("-", medianFU), medianFU]
x <- as.numeric(unlist(strsplit(x, "-")))
x1 <- x[seq(1, length(x), 2)]
x2 <- x[seq(2, length(x), 2)]
xbar <- rowMeans(cbind(x1, x2))
D <- D[grep("-", medianFU), medianFUNum := xbar]
D <- D[,
       `:=` (medianFUChar = medianFU,
             medianFU = medianFUNum,
             medianFUNum = NULL)]
unique(D[as.character(medianFU) != medianFUChar, .(medianFU, medianFUChar)])
D <- D[, medianFUChar := NULL]
```

Recode mal* values of `NR` and `-` to `NA`.


```{r}
recode <- function(x) {
    missval <- c("-", "NR", " ")
    vec <- D[, get(x)]
    vec[vec %in% missval] <- NA
    vec
}
col <- grep("^mal", names(D), value = TRUE)
for (i in 1:length(col)) {
    D[, col[i]] <- recode(col[i])
}
```

Substitute non-numeric text in mal* values.


```{r}
substitute <- function(x) {
    vec <- D[, get(x)]
    vec <- gsub("[[:alpha:]]", "", vec)
    vec
}
col <- grep("^mal", names(D), value = TRUE)
for (i in 1:4) {
    # Only substitute a subset of mal* columns
    D[, col[i]] <- substitute(col[i])
}
```

Convert mal* values to numeric.


```{r}
for (i in 1:4) {
    # Only convert a subset of mal* columns
    D <- D[, col[i] := as.numeric(get(col[i]))]
}
```

The `malAMLOrMDS` column, as entered, captures number of AML or MDS malignancies if the study reported them grouped as opposed to separately.
If the study reported them as separately, then these counts are captures in the `malAML` and `malMDS` columns.
I.e., the `malAML`, `malMDS`, and `malAMLOrMDS` columns are mutually exclusive counts.
We want to create another column, `malAMLOrMDSTotal`, to be non-mutually exclusive from `malAML`, `malMDS`, and `malAMLOrMDS`.
If none of the `malAML`, `malMDS`, and `malAMLOrMDS` values are populated, then set `malAMLOrMDSTotal` to `NA`, also.
NOTE: Couldn't get this to work: [http://stackoverflow.com/a/16513949/1427069](http://stackoverflow.com/a/16513949/1427069).


```{r}
s <- sample(D[, rowid], 12)
malAMLOrMDSTotal <- rowSums(D[, .(malAML, malMDS, malAMLOrMDS)], na.rm=TRUE)
D <- D[, malAMLOrMDSTotal := malAMLOrMDSTotal]
D <- D[is.na(malAML) & is.na(malMDS) & is.na(malAMLOrMDS), malAMLOrMDSTotal := NA]
D[s, .(rowid, malAML, malMDS, malAMLOrMDS, malAMLOrMDSTotal)]
```

Remove text from `nITT` column.


```{r}
D <- D[authorYear == "Misset (1996)" & arm == 2, nITT := "137"]
D <- D[authorYear == "Fumoleu (2003)" & arm == 1, nITT := "210"]
D <- D[authorYear == "Fumoleu (2003)" & arm == 2, nITT := "197"]
D <- D[authorYear == "Fumoleu (2003)" & arm == 3, nITT := "195"]
D <- D[, nITT := as.integer(nITT)]
```

Clean up the `regimen` and `dose` columns.
Output to [regimens.md](regimens.md) for checking.


```{r}
D <- D[, regimen := gsub("\\s$", "", regimen)]
D <- D[, regimen := gsub("\\r*\\n", " ", regimen)]
doseVar <- grep("dose", names(D), ignore.case = TRUE, value = TRUE)
for (i in 1:length(doseVar)) {
    D <- D[, doseVar[i] := gsub("\\r*\\n", " ", get(doseVar[i]))]
}
cat(kable(D[, .N, c("regimen", doseVar), with = TRUE][order(regimen)]),
    file = "regimens.md",
    sep = "\n")
file.info("regimens.md")
```

Calculate cumulative dose: $\text{total dose} \times \text{number of courses}$.


```{r}
D <- D[, anthracyclineTotalDose := as.numeric(anthracyclineTotalDose)]
D <- D[authorYear == "Bergh (2000)" & regimen == "FEC\nTailored", anthracyclineTotalDose := 75]
D <- D[authorYear == "Henderson (2003)", anthracyclineTotalDose := 75]
D <- D[authorYear == "Colleoni (2009)" & regimen == "AC-CMF", anthracyclineTotalDose := 75]
D <- D[, cyclophosphamideDose := as.numeric(cyclophosphamideDose)]
D <- D[authorYear == "Colleoni (2009)" & regimen == "EC", cyclophosphamideDose := 600]
D <- D[authorYear == "Bergh (2000)" & regimen == "FEC\nTailored", cyclophosphamideDose := 900]
D <- D[authorYear == "Henderson (2003)", anthracyclineTotalDose := 75]
D <- D[, taxaneCourses := as.numeric(taxaneCourses)]
D <- D[authorYear == "Shulman (2014)" & regimen == "T" & arm == 3, taxaneCourses := 4]
D <- D[authorYear == "Shulman (2014)" & regimen == "T" & arm == 4, taxaneCourses := 6]
D <- D[, fluoroucilTotalDose := as.numeric(fluoroucilTotalDose)]
D <- D[authorYear == "Bergh (2000)" & regimen == "FEC\nTailored", fluoroucilTotalDose := 600]
D <- D[authorYear == "Joensuu (2012)" & regimen == "TX-CEX", fluoroucilTotalDose := NA]
D <- D[!is.na(anthracyclineTotalDose) & !is.na(anthracyclineTotalDose),
       anthracyclineCumulDose := anthracyclineTotalDose * anthracyclineCourses]
D <- D[!is.na(cyclophosphamideDose) & !is.na(cyclophosphamideCourses),
       cyclophosphamideCumulDose := cyclophosphamideDose * cyclophosphamideCourses]
D <- D[!is.na(as.numeric(taxaneTotalDose)) & !is.na(taxaneCourses),
       taxaneCumulDose := as.numeric(taxaneTotalDose) * taxaneCourses]
D <- D[!is.na(fluoroucilTotalDose) & !is.na(fluoroucilCourses),
       fluoroucilCumulDose := fluoroucilTotalDose * fluoroucilCourses]
# D <- D[is.na(anthracyclineCumulDose), anthracyclineCumulDose := 0]
# D <- D[is.na(cyclophosphamideCumulDose), cyclophosphamideCumulDose := 0]
# D <- D[is.na(taxaneCumulDose), taxaneCumulDose := 0]
# D <- D[is.na(fluoroucilCumulDose), fluoroucilCumulDose := 0]
```

Check.


```{r}
head(D[!is.na(anthracyclineCumulDose),
       .(authorYear, regimen, anthracyclineCumulDose, anthracyclineTotalDose, anthracyclineCourses)])
head(D[!is.na(cyclophosphamideCumulDose),
       .(authorYear, regimen, cyclophosphamideCumulDose, cyclophosphamideDose, cyclophosphamideCourses)])
head(D[!is.na(taxaneCumulDose),
       .(authorYear, regimen, taxaneCumulDose, taxaneTotalDose, taxaneCourses)])
head(D[!is.na(fluoroucilCumulDose),
       .(authorYear, regimen, fluoroucilCumulDose, fluoroucilTotalDose, fluoroucilCourses)])
```

## Manual tidying

After an early round of analysis, we realized there were some `nITT` data entry errors in the commit [26115da](https://github.com/benjamin-chan/SecMalAfterBreastCaACT/raw/26115dab1eb2b075d40f545514a03c3b30cde6b4/SecondaryMalignancies_Data_BG_3_4.xlsx) (Jun 1, 2016) spreadsheet.


```{r}
D <- D[authorYear == "Mamounas (2005)" & nITT == 1528, `:=` (nITTOld = nITT, nITT = 1529, isFixed = TRUE)]
D <- D[authorYear == "Fumoleu (2003)" & nITT == 210, `:=` (nITTOld = nITT, nITT = 212, isFixed = TRUE)]
D <- D[authorYear == "Fumoleu (2003)" & nITT == 197, `:=` (nITTOld = nITT, nITT = 209, isFixed = TRUE)]
D <- D[authorYear == "Fumoleu (2003)" & nITT == 195, `:=` (nITTOld = nITT, nITT = 200, isFixed = TRUE)]
D <- D[authorYear == "Fargeot (2004)" & nITT == 155, `:=` (nITTOld = nITT, nITT = 164, isFixed = TRUE)]
D <- D[authorYear == "Fargeot (2004)" & nITT == 163, `:=` (nITTOld = nITT, nITT = 174, isFixed = TRUE)]
D <- D[authorYear == "Kerbrat (2007)" & nITT == 235, `:=` (nITTOld = nITT, nITT = 241, isFixed = TRUE)]
D <- D[authorYear == "Kerbrat (2007)" & nITT == 236, `:=` (nITTOld = nITT, nITT = 241, isFixed = TRUE)]
D <- D[authorYear == "Fisher (1997)" & nITT == 763, `:=` (nITTOld = nITT, nITT = 767, isFixed = TRUE)]
D <- D[authorYear == "Fisher (1997)" & nITT == 771, `:=` (nITTOld = nITT, nITT = 772, isFixed = TRUE)]
D <- D[authorYear == "Bonneterre (2005)" & nITT == 271, `:=` (nITTOld = nITT, nITT = 289, isFixed = TRUE)]
D <- D[authorYear == "Bonneterre (2005)" & nITT == 266, `:=` (nITTOld = nITT, nITT = 276, isFixed = TRUE)]
D <- D[authorYear == "Roche (2006)" & nITT == 163, `:=` (nITTOld = nITT, nITT = 164, isFixed = TRUE)]
D <- D[authorYear == "Roche (2006)" & nITT == 168, `:=` (nITTOld = nITT, nITT = 169, isFixed = TRUE)]
D <- D[authorYear == "Shulman (2014)" & nITT == 1107, `:=` (nITTOld = nITT, nITT = 1142, isFixed = TRUE)]
D <- D[authorYear == "Shulman (2014)" & nITT ==  766, `:=` (nITTOld = nITT, nITT =  789, isFixed = TRUE)]
D <- D[authorYear == "Shulman (2014)" & nITT == 1119, `:=` (nITTOld = nITT, nITT = 1151, isFixed = TRUE)]
D <- D[authorYear == "Shulman (2014)" & nITT ==  782, `:=` (nITTOld = nITT, nITT =  789, isFixed = TRUE)]
D <- D[authorYear == "Swain (2013)" & nITT == 1617, `:=` (nITTOld = nITT, nITT = 1630, isFixed = TRUE)]
D <- D[authorYear == "Swain (2013)" & nITT == 1624, `:=` (nITTOld = nITT, nITT = 1634, isFixed = TRUE)]
D <- D[authorYear == "Swain (2013)" & nITT == 1618, `:=` (nITTOld = nITT, nITT = 1630, isFixed = TRUE)]
D[isFixed == TRUE, .(authorYear, regimen, nITTOld, nITT)]
D <- D[, `:=` (nITTOld = NULL, isFixed = NULL)]
```

We also made the decision to use the AML/MDS outcomes reported in Perez (2011) as the outcomes for Romond (2005) since they study the same cohort. Other edits to *AML or MDS* and *non-breast solid* malignancy outcome were needed, as well.


```{r}
D <- D[authorYear == "Martin (2010)" & regimen == "TAC",
      `:=` (malAMLOrMDSTotalOld = malAMLOrMDSTotal, malAMLOrMDSTotal = 0, isFixed = TRUE)]
D <- D[authorYear == "Martin (2010)" & regimen == "FAC",
      `:=` (malAMLOrMDSTotalOld = malAMLOrMDSTotal, malAMLOrMDSTotal = 0, isFixed = TRUE)]
D <- D[authorYear == "Romond (2005)" & regimen == "AC-T",
      `:=` (malAMLOrMDSTotalOld = malAMLOrMDSTotal, malAMLOrMDSTotal = 2, isFixed = TRUE)]
D <- D[authorYear == "Romond (2005)" & regimen == "ACT-T-Trast",
      `:=` (malAMLOrMDSTotalOld = malAMLOrMDSTotal, malAMLOrMDSTotal = 1, isFixed = TRUE)]
D <- D[authorYear == "Del Mastro (2015)",
      `:=` (malAMLOrMDSTotalOld = malAMLOrMDSTotal, malAMLOrMDSTotal = 0, isFixed = TRUE)]
D <- D[authorYear == "Del Mastro (2015)" & regimen == "EC-T" & nITT == 502,
      `:=` (malAMLOrMDSTotalOld = malAMLOrMDSTotal, malAMLOrMDSTotal = 2, isFixed = TRUE)]

D <- D[authorYear == "Bernard-Marty (2003)" & regimen == "EC" & nITT == 267,
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 4, isFixed = TRUE)]
D <- D[authorYear == "Bonneterre (2005)" & regimen == "FEC 50",
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 11, isFixed = TRUE)]
D <- D[authorYear == "Bonneterre (2005)" & regimen == "FEC 100",
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 15, isFixed = TRUE)]
D <- D[authorYear == "Citron (2003)" & regimen == "A-T-C" & nITT == 493,
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 11, isFixed = TRUE)]
D <- D[authorYear == "Citron (2003)" & regimen == "AC-T" & nITT == 495,
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 11, isFixed = TRUE)]
D <- D[authorYear == "Eiermann (2011)" & regimen == "AC-T",
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 34, isFixed = TRUE)]
D <- D[authorYear == "Eiermann (2011)" & regimen == "TAC",
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 31, isFixed = TRUE)]
D <- D[authorYear == "Romond (2005)" & regimen == "AC-T",
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 13, isFixed = TRUE)]
D <- D[authorYear == "Romond (2005)" & regimen == "ACT-T-Trast",
      `:=` (malNonBreastSolidOld = malNonBreastSolid, malNonBreastSolid = 4, isFixed = TRUE)]

D[isFixed == TRUE,
  .(authorYear, regimen, nITT, malAMLOrMDSTotalOld, malAMLOrMDSTotal, malNonBreastSolidOld, malNonBreastSolid)]
D <- D[, `:=` (malAMLOrMDSTotalOld = NULL, malNonBreastSolidOld = NULL, isFixed = NULL)]
```

Deduplicate.


```{r}
n0 <- nrow(D)
setkey(D, authorYear, trial, arm)
D <- unique(D)
message(sprintf("Removed %d duplicate row(s) (%.03g%%)", n0 - nrow(D), (n0 - nrow(D)) / n0 * 100))
```


Save data.table to file.

```{r}
f <- "data.RData"
save(D, file = f)
file.info(f)[c("size", "mtime")]
```
